---
title: "Attrition Case Study"
author: "Zach Harris"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
# Welcome to the Attrition Case Study
The purpose of this case study aims to examine and study the data set to determine two key items:
 
 1. Determine which variables, if any, are able to predict the attrition of an employee
 2. If variables within the data set are able to predict the Monthly Income for an employee
 
In order to conduct this study, the data had to be examined and cleaned.
First, an analysis was performed to determine if there were any missing values in the data set.


 
```{r Libraries and Packages, echo = FALSE}
#Call Libraries#
library(dplyr)
library(ggplot2)
library(dlookr)
library(visdat)
library(GGally)
library(forcats)
library(class)
library(caret)
library(e1071)
library(magrittr)
library(leaps)
library(ggthemes)
```

```{r NA Data, echo=FALSE}
caseStudy<-data.frame(read.csv('/Users/ZacharyHarris/Desktop/MS Data Science/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2-data.csv'))
vis_miss(caseStudy) #NA Values
```

From the graph above, it can be seen that there were no missing values from the data set. We are also able to see the amount of columns
or variables that the data set has. Now that we are sure the data is clean and ready, we are ready to begin anaylsis and manipulation.

## Exploratory Data Analysis
To begin the analysis, we began by searching for relationships between variables. Some examples are shown below. The accompanying code is provided from customer-consultant transparency. 
  

```{r Job Roles, echo = TRUE}

#Job Role Histogram
ggplot(caseStudy ,aes(x=fct_infreq(JobRole), fill = JobRole)) + 
  geom_histogram(stat = 'count')+ 
  coord_flip() + theme_clean() + ggtitle('Number of Employees by Job Role')+
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(stat="count", aes(label=..count.., hjust =-0.5, vjust=0)) + 
  xlab('Job Role') + ylab('Number of Employees')

#Job Role Histogram with Attrition
ggplot(caseStudy ,aes(x=fct_infreq(JobRole), fill = Attrition)) + 
  geom_histogram(stat = 'count', position = 'dodge')+ 
  coord_flip() + theme_clean() + ggtitle('Number of Employees by Job Role')+
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(stat="count", aes(label=..count.., hjust =-0.5, vjust=0)) + 
  xlab('Job Role') + ylab('Number of Employees')

#Monthly Income by Education Field
ggplot(caseStudy, aes(x=fct_infreq(EducationField), y = MonthlyIncome, fill = Attrition)) +
  geom_boxplot() +theme_clean() + xlab('Education Field') +ylab('Monthly Income') + 
  ggtitle('Incomes by Education Field') + theme_clean() + 
  theme(plot.title = element_text(hjust = 0.5)) 

#Attrition Histogram
ggplot(caseStudy, aes(x=fct_infreq(Attrition), fill = Attrition)) +
  geom_histogram(stat = "count") + ylim(0,800) +
  ggtitle('Number of Attritions') + theme_clean() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0))+
  xlab('Attrition (Yes/No)')+ ylab('Number of Employees')

#Job Satisfaction and Attrition
ggplot(caseStudy, aes(x=JobSatisfaction, fill = Attrition)) + theme_clean()+
  geom_bar(stat="count", position = "dodge") + 
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0)) +
  xlab('Job Satisfaction Rating') + ylab('Number of Employees') + 
  ggtitle('Job Satisfaction Ratings')+ theme(plot.title = element_text(hjust = 0.5))

#Relationship Satisfaction and Attrition
ggplot(caseStudy, aes(x=RelationshipSatisfaction, fill = Attrition)) + 
  geom_bar(stat="count", position = "dodge") +
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0)) +theme_clean()+
  xlab('Relationship Satisfaction Rating') + ylab('Number of Employees') + 
  ggtitle('Relationship Satisfaction Ratings')+ theme(plot.title = element_text(hjust = 0.5)) 

#Environment Satisfaction and Attrition
ggplot(caseStudy, aes(x=EnvironmentSatisfaction, fill = Attrition)) + 
  geom_bar(stat="count", position = "dodge") +
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0))+theme_clean()+
  xlab('Environment Satisfaction Rating') + ylab('Number of Employees') + 
  ggtitle('Environment Satisfaction Ratings')+ theme(plot.title = element_text(hjust = 0.5)) 

#Job Involvement and Attrition
ggplot(caseStudy, aes(x=JobInvolvement, fill = Attrition)) + theme_clean()+
  geom_bar(stat="count", position = "dodge") +
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0))+
  xlab('Job Involvement Rating') + ylab('Number of Employees') + 
  ggtitle('Job Involvement Ratings')+ theme(plot.title = element_text(hjust = 0.5)) 

#Creating Overall Satisfaction variable
caseStudy$OverallSatisfaction = (caseStudy$JobSatisfaction+
                                   caseStudy$RelationshipSatisfaction+
                                   caseStudy$EnvironmentSatisfaction+
                                   caseStudy$JobInvolvement)

#Overall Satisfaction and Attrition
ggplot(caseStudy, aes(x=OverallSatisfaction, fill = Attrition)) + theme_clean()+
  geom_bar(stat="count", position = "dodge")  +
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0)) + 
  xlab('Overall Satisfaction') + ylab('Number of Employees') + 
  ggtitle('Overall Satisfaction Rating')+ theme(plot.title = element_text(hjust = 0.5))

#Age and Attrition
ggplot(caseStudy, aes(x=Age, fill = Attrition)) + theme_clean()+
  geom_bar(stat="count", position = "dodge") +
  geom_text(stat="count", aes(label=..count.., hjust =0.5, vjust=-1.0))+
  xlab('Age') + ylab('Number of Employees') + 
  ggtitle('Attrition by Age')+ theme(plot.title = element_text(hjust = 0.5)) 


#Stock Option Level and Attrition
ggplot(caseStudy, aes (x=StockOptionLevel, fill = Attrition)) + theme_clean()+
  geom_bar(stat = "count", position = "dodge") + theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(stat = "count", aes(label=..count.., hjust = 0.5, vjust = -1.0)) +
  xlab('Stock Option Level') + ylab('Number of Employees') + ggtitle('Stock Option Levels')

#Age and Monthly Income
ggplot(caseStudy, aes(x= Age, y=MonthlyIncome, color = Attrition))+ theme_clean()+
  geom_point() + xlab('Age') + ylab('Monthly Income') + ggtitle('Monthly Income vs. Age') +
  theme(plot.title = element_text(hjust = 0.5))

#Job Role and Monthly Income
ggplot(caseStudy, aes(x= JobRole, y=MonthlyIncome, color = Attrition))+
  geom_boxplot() + theme_clean() + xlab('Job Role') + ylab('Monthly Income') + 
  ggtitle('Job Role vs. Age') +
  theme(plot.title = element_text(hjust = 0.5))
```  



```{r Correlations}
# Correlations
ggpairs(select(caseStudy, JobLevel, MonthlyIncome, TotalWorkingYears,
               YearsInCurrentRole, YearsAtCompany,PerformanceRating,
               PercentSalaryHike,YearsWithCurrManager,YearsSinceLastPromotion))
```

## kNN Classifier
```{r kNN Classifier}

#Split data set into partially test and partially train 
splitPerc = .70     

#Filters data set, unneeded since all is yes and no. 
#If the attrition value for a specific row was left blank, this would eliminate the probelm
attritionYesNo = caseStudy %>% filter(Attrition == "Yes" | Attrition == "No")
summary(attritionYesNo)

#Assining test and train
trainIndices = sample(1:dim(attritionYesNo)[1],round(splitPerc * dim(attritionYesNo)[1]))
train = attritionYesNo[trainIndices,]
test = attritionYesNo[-trainIndices,]

# k = ?
classifications = knn(train[,c(16,30,37)],test[,c(16,30,37)],train$Attrition, prob = TRUE, k = 13)
table(classifications,test$Attrition)
confusionMatrix(table(classifications,test$Attrition))

testNoAttrition<-data.frame(read.csv('/Users/ZacharyHarris/Desktop/MS Data Science/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Attrition.csv'))

#Creating Overall Satisfaction variable
testNoAttrition$OverallSatisfaction = (testNoAttrition$JobSatisfaction+
                                         testNoAttrition$RelationshipSatisfaction+
                                         testNoAttrition$EnvironmentSatisfaction+
                                         testNoAttrition$JobInvolvement)

prediction = knn(train[1:300,c(14,28,36)],testNoAttrition[,c(14,28,36)],train[1:300,c(3)], prob = TRUE, k = 13)
table(prediction,test$Attrition[1:300])
confusionMatrix(table(prediction,test$Attrition[1:300]))

prediction_final<-data.frame(testNoAttrition$ID,prediction)
colnames(prediction_final)<-c('ID','Attrition')
write.csv(prediction_final,"/Users/ZacharyHarris/Desktop/Case2PredictionsMissingAttrition.csv", row.names = FALSE)

# Loop for many k and one training / test partition

accs = data.frame(accuracy = numeric(30), k = numeric(30))
sens = data.frame(sensitivity = numeric(30), k = numeric(30))
spec = data.frame(specificity = numeric(30), k = numeric(30))

for(i in 1:30)
{
  classifications = knn(train[,c(16,30,37)],test[,c(16,30,37)],train$Attrition, prob = TRUE, k = i)
  table(test$Attrition,classifications)
  CM = confusionMatrix(table(test$Attrition,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
  sens$sensitivity[i]=CM$byClass[1]
  sens$k[i] = i
  spec$specificity[i]=CM$byClass[2]
  spec$k[i] = i
}
par(mfrow=c(3,1))
plot(accs$k,accs$accuracy, type = "l", xlab = "k")
axis(1, xaxp=c(0, 30, 30))
plot(sens$k,sens$sensitivity, type ="l", xlab = "k")
axis(1, xaxp=c(0, 30, 30))
plot(spec$k,spec$specificity, type = "l", xlab = "k")
axis(1, xaxp=c(0, 30, 30))
```

##Linear Regression
```{r Linear Regression}
#####Linear Regression#####


#Modify caseStudy dataframe to contain on/off switches for qualitative variables
#Eliminating columns I do not need for the next set of operations
caseStudyModified <- subset(caseStudy,select=-c(ID, Over18, Attrition))

#BusinessTravel - 
caseStudyModified$BusinessTravel[caseStudyModified$BusinessTravel=="Non-Travel"]<-0
caseStudyModified$BusinessTravel[caseStudyModified$BusinessTravel=="Travel_Frequently"]<-1
caseStudyModified$BusinessTravel[caseStudyModified$BusinessTravel=="Travel_Rarely"]<-2

#Department - 
caseStudyModified$Department[caseStudyModified$Department=="Sales"]<-0
caseStudyModified$Department[caseStudyModified$Department=="Human Resources"]<-1
caseStudyModified$Department[caseStudyModified$Department=="Research & Development"]<-2

#EducationField - 
caseStudyModified$EducationField[caseStudyModified$EducationField=="Human Resources"]<-0
caseStudyModified$EducationField[caseStudyModified$EducationField=="Technical Degree"]<-1
caseStudyModified$EducationField[caseStudyModified$EducationField=="Medical"]<-2
caseStudyModified$EducationField[caseStudyModified$EducationField=="Life Sciences"]<-3
caseStudyModified$EducationField[caseStudyModified$EducationField=="Other"]<-4
caseStudyModified$EducationField[caseStudyModified$EducationField=="Marketing"]<-5

#Gender - 
caseStudyModified$Gender[caseStudyModified$Gender=="Female"]<-0
caseStudyModified$Gender[caseStudyModified$Gender=="Male"]<-1

#JobRole - 
caseStudyModified$JobRole[caseStudyModified$JobRole=="Sales Representative"]<-0
caseStudyModified$JobRole[caseStudyModified$JobRole=="Human Resources"]<-1
caseStudyModified$JobRole[caseStudyModified$JobRole=="Lab Technician"]<-2
caseStudyModified$JobRole[caseStudyModified$JobRole=="Laboratory Technician"]<-2
caseStudyModified$JobRole[caseStudyModified$JobRole=="Research Scientist"]<-3
caseStudyModified$JobRole[caseStudyModified$JobRole=="Sales Executive"]<-4
caseStudyModified$JobRole[caseStudyModified$JobRole=="Manufacturing Director"]<-5
caseStudyModified$JobRole[caseStudyModified$JobRole=="Healthcare Representative"]<-6
caseStudyModified$JobRole[caseStudyModified$JobRole=="Research Director"]<-7
caseStudyModified$JobRole[caseStudyModified$JobRole=="Manager"]<-8

#MaritalStatus - 
caseStudyModified$MaritalStatus[caseStudyModified$MaritalStatus=="Divorced"]<-0
caseStudyModified$MaritalStatus[caseStudyModified$MaritalStatus=="Single"]<-1
caseStudyModified$MaritalStatus[caseStudyModified$MaritalStatus=="Married"]<-2

#OverTime - 
caseStudyModified$OverTime[caseStudyModified$OverTime=="No"]<-0
caseStudyModified$OverTime[caseStudyModified$OverTime=="Yes"]<-1



#Determine fit using all variables. Elimination will be based on whether variable has effect on model

fit = lm(MonthlyIncome~BusinessTravel + DailyRate + Department + DistanceFromHome + 
           Education + EducationField + EmployeeCount + EmployeeNumber + 
           EnvironmentSatisfaction + Gender + HourlyRate + JobInvolvement + 
           JobLevel + JobRole + JobSatisfaction + MaritalStatus + MonthlyIncome + 
           MonthlyRate + NumCompaniesWorked + OverTime + PercentSalaryHike + 
           PerformanceRating + RelationshipSatisfaction + StandardHours + 
           StockOptionLevel + TotalWorkingYears + TrainingTimesLastYear + 
           WorkLifeBalance + YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + 
           YearsWithCurrManager, caseStudyModified)
summary(fit)


#Eliminate based on level of effect on model
fit = lm(MonthlyIncome~BusinessTravel + 
           JobLevel + JobRole  + PerformanceRating  + YearsSinceLastPromotion, caseStudyModified)
summary(fit)


#Eliminated a second time based on level of effect on model
fit = lm(MonthlyIncome~BusinessTravel + 
           JobLevel + JobRole  + YearsSinceLastPromotion, caseStudyModified)
summary(fit)
mse<-mean(residuals(fit)^2)
rmse<-sqrt(mse)
rmse
prediction = data.frame(caseStudy$ID,predict(fit))
colnames(prediction)<-c('ID','MonthlyIncome')
prediction
write.csv(prediction,"/Users/ZacharyHarris/Desktop/untitled folder\\Case2PredictionsXXXXAttrition.csv", row.names = FALSE)
#Line above will save predictions based on the values that were on the original dataset
#For example, if you want to see what it estimated from your own set, this would be those values.




#Loading dataset without salary values to test the model
nosalary<-read.csv('/Users/ZacharyHarris/Desktop/MS Data Science/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Salary.csv')

#Preparing data set to work with my parameters. We made qualitative variables switches.
#So, we have to do the same thing here.

#BusinessTravel - 
nosalary$BusinessTravel[nosalary$BusinessTravel=="Non-Travel"]<-0
nosalary$BusinessTravel[nosalary$BusinessTravel=="Travel_Frequently"]<-1
nosalary$BusinessTravel[nosalary$BusinessTravel=="Travel_Rarely"]<-2

#Department - 
nosalary$Department[nosalary$Department=="Sales"]<-0
nosalary$Department[nosalary$Department=="Human Resources"]<-1
nosalary$Department[nosalary$Department=="Research & Development"]<-2

#EducationField - 
nosalary$EducationField[nosalary$EducationField=="Human Resources"]<-0
nosalary$EducationField[nosalary$EducationField=="Technical Degree"]<-1
nosalary$EducationField[nosalary$EducationField=="Medical"]<-2
nosalary$EducationField[nosalary$EducationField=="Life Sciences"]<-3
nosalary$EducationField[nosalary$EducationField=="Other"]<-4
nosalary$EducationField[nosalary$EducationField=="Marketing"]<-5

#Gender - 
nosalary$Gender[nosalary$Gender=="Female"]<-0
nosalary$Gender[nosalary$Gender=="Male"]<-1

#JobRole - 
nosalary$JobRole[nosalary$JobRole=="Sales Representative"]<-0
nosalary$JobRole[nosalary$JobRole=="Human Resources"]<-1
nosalary$JobRole[nosalary$JobRole=="Lab Technician"]<-2
nosalary$JobRole[nosalary$JobRole=="Laboratory Technician"]<-2
nosalary$JobRole[nosalary$JobRole=="Research Scientist"]<-3
nosalary$JobRole[nosalary$JobRole=="Sales Executive"]<-4
nosalary$JobRole[nosalary$JobRole=="Manufacturing Director"]<-5
nosalary$JobRole[nosalary$JobRole=="Healthcare Representative"]<-6
nosalary$JobRole[nosalary$JobRole=="Research Director"]<-7
nosalary$JobRole[nosalary$JobRole=="Manager"]<-8

#MaritalStatus - 
nosalary$MaritalStatus[nosalary$MaritalStatus=="Divorced"]<-0
nosalary$MaritalStatus[nosalary$MaritalStatus=="Single"]<-1
nosalary$MaritalStatus[nosalary$MaritalStatus=="Married"]<-2

#OverTime - 
nosalary$OverTime[nosalary$OverTime=="No"]<-0
nosalary$OverTime[nosalary$OverTime=="Yes"]<-1


#Using model to predict monthly income
prediction_final = data.frame(nosalary$ID,predict(fit, newdata = nosalary))
colnames(prediction_final)<-c('ID','MonthlySalary')
write.csv(prediction_final,"/Users/ZacharyHarris/Desktop/ZacharyHarris_Case2PredictionsMonthlyIncome.csv", row.names = FALSE)
```


